Using an iterated block particle filter via \pkg{spatPomp}

### 2022-12-12 bm.Rnw
ready for testing on gl

%%%%%% iiiiiiii %%%%
<<bm-model,echo=T,eval=T>>=
library(spatPomp)
i <- 2
b <- bm2(U=4,N=switch(i,20,200),unit_specific_names="rho")
plot(b)
@

@@

<<m-prof-sig,eval=T,echo=F>>=
stew(file=paste0(out_dir,"m_prof_R01.rda"),seed=722,{
m_prof_R0_init <- rep(30,U)
names(m_prof_R0_init) <- paste0("R0",1:U)
m_prof_R0_vals <- seq(from=15,to=45,length=switch(i,5,20))
tic <- Sys.time()
m_R0 <- m
u <- 2
R0_start <- seq(from=15,to=45,length=U)
params_R0_start <- coef(m_R0)
params_R0_start[paste0("R0",1:U)] <- R0_start
reps <- switch(i,3,10)
param_R0_start_matrix <- matrix(params_R0_start,byrow=TRUE,
  ncol=length(coef(m_R0)),nrow=reps*length(m_prof_R0_vals),
  dimnames=list(NULL,names(coef(m_R0))))
param_R0_start_matrix[,paste0("R0",u)] <- rep(m_prof_R0_vals,each=reps)
m_prof_R0_searches <- foreach(s=1:nrow(param_R0_start_matrix))%dopar%{
  rw.sd_call <- rep(list(0.02),times=U-1)
  names(rw.sd_call) <- lapply((1:U)[-u],function(x)paste0("R0",x))
  ibpf(m_R0,
    params=param_R0_start_matrix[s,],
    Nbpf=switch(i,2,50),Np=switch(i,20,2000),
    rw.sd=do.call(rw.sd,rw.sd_call),
    unitParNames="R0",
    sharedParNames=NULL,
    block_size=1,
    spat_regression=0,
    cooling.fraction.50=0.5
  )
}
toc <- Sys.time()
})
m_prof_R02time <- toc-tic
@

<<m-prof-sig-lik,eval=T,echo=F>>=
stew(file=paste0(out_dir,"m_prof_R0",u,"_lik.rda"),seed=868,{
tic <- Sys.time()
m_bpf_R0_eval <- foreach(mfit=m_prof_R0_searches,.combine=cbind) %dopar% {
  replicate(switch(i,5,10),
    logLik(bpfilter(mfit,block_size=1,Np=switch(i,20,5000)))
  )
}
m_pf_R0_eval <- foreach(mfit=m_prof_R0_searches,.combine=cbind) %dopar% {
  replicate(switch(i,5,10),
    logLik(pfilter(mfit,Np=switch(i,20,5000)))
  )
}
toc <- Sys.time()
})
m_prof_R02eval_time <- toc-tic
@

The $\R0_1$ profile took \Sexpr{myround(as.double(m_prof_R02time,units="mins"),2)} mins using \Sexpr{cores} computing cores.
Likelihood evaluation took \Sexpr{myround(as.double(m_prof_R02eval_time,units="mins"),2)}, shared between BPF and PF.

<<m-prof-sig1-plot, fig.height=8, fig.width=6, out.width="5.5in", fig.cap = "Top: profile for $R0_1$ using an IBPF search with likelihood computed using BPF, for $K=4$ blocks each having 1 unit. Middle: Exact profile (dashed red line) and the same IBPF search with likelihood computed exactly using the Kalman filter. Bottom: The same IBPF search with likelihood computed using the particle filter.",eval=T,echo=F>>=

@@

##### window to avoid problematic particle filter behavior at extremes
prof_include <- (m_prof_R0_vals > 20) & (m_prof_R0_vals < 40)

##### bpf evaluation
m_p2_R0_bpf <- apply(matrix(apply(m_bpf_R0_eval,2,mean),nrow=reps),2,max)[prof_include]
bpf_prof_R0_range <- c(max(m_p2_R0_bpf)-80,max(m_p2_R0_bpf))
plot(y=m_p2_R0_bpf,x=m_prof_R0_vals[prof_include], ylim=bpf_prof_R0_range,
  ylab="Log-likelihood")
m_p2_R0_bpf_mcap <- mcap(logLik=m_p2_R0_bpf,
  parameter=m_prof_R0_vals[prof_include])
lines(m_p2_R0_bpf_mcap$fit$parameter,m_p2_R0_bpf_mcap$fit$smoothed)

>>>>

%%%%%% iiiiiiii %%%%
<<bm-model,echo=T,eval=T>>=
library(spatPomp)
i <- 1
b <- bm2(U=4,N=switch(i,20,200),unit_specific_names="rho")
plot(b)
@

@@

<<m-prof-R0,eval=T,echo=F>>=
stew(file=paste0(out_dir,"m_prof_R02.rda"),seed=722,{
m_prof_R0_init <- rep(30,U)
names(m_prof_R0_init) <- paste0("R0",1:U)
m_prof_R0_vals <- seq(from=25,to=35,length=switch(i,5,20))
tic <- Sys.time()
m_R0 <- m
u <- 2
R0_start <- seq(from=25,to=35,length=U)
params_R0_start <- coef(m_R0)
params_R0_start[paste0("R0",1:U)] <- R0_start
reps <- switch(i,3,10)
param_R0_start_matrix <- matrix(params_R0_start,byrow=TRUE,
  ncol=length(coef(m_R0)),nrow=reps*length(m_prof_R0_vals),
  dimnames=list(NULL,names(coef(m_R0))))
param_R0_start_matrix[,paste0("R0",u)] <- rep(m_prof_R0_vals,each=reps)
m_prof_R0_searches <- foreach(s=1:nrow(param_R0_start_matrix))%dopar%{
  rw.sd_call <- rep(list(0.02),times=U-1)
  names(rw.sd_call) <- lapply((1:U)[-u],function(x)paste0("R0",x))
  ibpf(m_R0,
    params=param_R0_start_matrix[s,],
    Nbpf=switch(i,2,50),Np=switch(i,20,2000),
    rw.sd=do.call(rw.sd,rw.sd_call),
    unitParNames="R0",
    sharedParNames=NULL,
    block_size=1,
    spat_regression=0,
    cooling.fraction.50=0.5
  )
}
toc <- Sys.time()
})
m_prof_R02time <- toc-tic
@

<<m-prof-R0-lik,eval=T,echo=F>>=
stew(file=paste0(out_dir,"m_prof_R0",u,"_lik.rda"),seed=868,{
tic <- Sys.time()
m_bpf_R0_eval <- foreach(mfit=m_prof_R0_searches,.combine=cbind) %dopar% {
  replicate(switch(i,5,10),
    logLik(bpfilter(mfit,block_size=1,Np=switch(i,20,5000)))
  )
}
m_pf_R0_eval <- foreach(mfit=m_prof_R0_searches,.combine=cbind) %dopar% {
  replicate(switch(i,5,10),
    logLik(pfilter(mfit,Np=switch(i,20,5000)))
  )
}
toc <- Sys.time()
})
m_prof_R02eval_time <- toc-tic
@

The $R0_2$ profile took \Sexpr{myround(as.double(m_prof_R02time,units="mins"),2)} mins using \Sexpr{cores} computing cores.
Likelihood evaluation took \Sexpr{myround(as.double(m_prof_R02eval_time,units="mins"),2)}, shared between BPF and PF.

<<m-prof-R02-plot, fig.height=8, fig.width=6, out.width="5.5in", fig.cap = "Top: profile for $R0_2$ using an IBPF search with likelihood computed using BPF, for $K=4$ blocks each having 1 unit. Bottom: The same IBPF search with likelihood computed using the particle filter.",eval=T,echo=F>>=

@@

##### window if needed to avoid extremes with filter failures
prof_include <- (m_prof_R0_vals > 20) & (m_prof_R0_vals < 40)

##### bpf evaluation
m_p2_R0_bpf <- apply(matrix(apply(m_bpf_R0_eval,2,mean),nrow=reps),2,max)[prof_include]
bpf_prof_R0_range <- c(max(m_p2_R0_bpf)-80,max(m_p2_R0_bpf))
plot(y=m_p2_R0_bpf,x=m_prof_R0_vals[prof_include], ylim=bpf_prof_R0_range,
  ylab="Log-likelihood",xlab="")
m_p2_R0_bpf_mcap <- mcap(logLik=m_p2_R0_bpf,
  parameter=m_prof_R0_vals[prof_include])
lines(m_p2_R0_bpf_mcap$fit$parameter,m_p2_R0_bpf_mcap$fit$smoothed)
